---
title: "Advanced Econometrics 1 – PS3"
author: "Lukas Jürgensmeier"
date: "17 11 2019"
output: pdf_document
numbersections: true
---

# Fixed Effects (FE) vs. Random Effects (RE)

# Pooled OLS (POLS) and RE (Vella and Verbeek 1998)

see https://rdrr.io/cran/wooldridge/man/wagepan.html for variable description.

## Difference between Pooled OLS and Random Effects

Explain the difference between POLS and RE with respect to the underlying assumptions.

## Estimate Pooled OLS and Random Effects Model on Wagepan Data

Using log(wage) as dependent variable estimate a model that contains an intercept, the regressor union, the
year dummies d81 through d87, educ, black, hisp. Consider the POLS and RE estimation. How do coefficients
and standard errors compare?

```{r results = "hide", message = FALSE, warning = FALSE}
library(plm)
library(lmtest)
library(sandwich)
library(stargazer)
```

```{r}
wagepan <- read.csv("data/PS3_wagepan.csv")
```

```{r}
model_ols <- lm(lwage ~ union + d81 + d82 + d83 + d84 + d85 + d86 + d87 + educ + black + hisp,
                data = wagepan)
model_pols <- plm(lwage ~ union + d81 + d82 + d83 + d84 + d85 + d86 + d87 + educ + black + hisp,
                  model = "pooling",
                  index = c("nr", "year"),
                  data = wagepan)
model_re <- plm(lwage ~ union + d81 + d82 + d83 + d84 + d85 + d86 + d87 + educ + black + hisp,
                  model = "random",
                  index = c("nr", "year"),
                  data = wagepan)
```

Coefficients and standard errors are rather similar and there is no obvious pattern in their small differences

## Obtain the robust standard errors for POLS in your estimation. Why do we need them in the POLS estimation?

Followed this method:

https://stats.stackexchange.com/questions/145650/estimating-robust-standard-errors-in-panel-data-regressions

```{r}
pols_robust_se <- coeftest(model_pols, vcov = vcovHC(model_pols, type = "HC0")) # , cluster = "group" ?
```

## What happens if you estimate the equation by FE? Compare the estimates.

```{r}
model_fe <- plm(lwage ~ union + d81 + d82 + d83 + d84 + d85 + d86 + d87 + educ + black + hisp,
                  model = "within",
                  index = c("nr", "year"),
                  data = wagepan)
```

```{r results = "asis"}
stargazer(model_ols, model_pols, pols_robust_se, model_re, model_fe,
          title = "Model Comparison",
          omit.stat = c("f"),
          column.labels = c("OLS", "Pooled OLS", "POLS (robust SE)", "Random Effects", "Fixed Effects"),
          model.names = FALSE, header = FALSE, font.size = "small")
```